---
title: "Quarto Guide"
format:
  pdf:
    documentclass: scrartcl
    pdf-engine: xelatex
    code-block:
      wrap: true
    header-includes:
      - \usepackage[dvipsnames]{xcolor}
      - \usepackage{fontawesome5}
      - \usepackage{graphicx} % for scalebox
      - \usepackage{tcolorbox}
      - \tcbuselibrary{skins, breakable, hooks}
      - \usepackage{fontawesome5}
      - \usepackage{wrapfig}
      - \usepackage{lipsum} % just for filler text
      - \usepackage{array}
filters:
  - critic-markup
extensions:
  - quarto-ext/fontawesome
  - shafayetShafee/bsicons
---

# Quarto Overview

Quarto is a multi-language successor to `R Markdown` from Posit, the company that distributes RStudio. Like `R Markdown`, Quarto allows one to create dynamic documents with code, text, and outputs (HTML, Word, PDF) from a single source file, but it supports multiple languages beyond R (including Python and Julia). It uses the `knitr` engine to execute R code, so most existing `.rmd` files can be rendered without modification.

Quarto is separate, standalone software and not just an R package, which means R is not required for basic functionality. Unlike `R Markdown`, which depended on extra packages to extend its capabilities, Quarto integrates features such as figure cross-referencing, book creation, website/blog creation, and diagramming (Mermaid, Graphviz) directly.

In contrast to `R Markdown`, which used R-specific chunk options (e.g., `echo = FALSE`, `warning = FALSE`), Quarto adopts a unified YAML syntax for configuration in both the document header and individual code blocks. This consistency applies across languages (R, Python, and Julia), though some options are format-specific. For example, certain settings affect HTML output, while others apply only to PDF.

Beyond its built‑in features, Quarto supports **extensions** — modular add‑ons that contribute filters, themes, and tools for specialized tasks. Extensions make it easy to incorporate functionality such as icon sets, diagramming filters, or markup processors without writing custom code. They are installed at the project level, which keeps them version‑controlled alongside your source files and ensures that collaborators can render documents consistently. The guide to installation will overview how to install extensions after discussing how to install the Quarto software.

## Installing and Verifying Quarto

1. **Install Quarto**
  - Go to [Getting Started With Quarto](https://quarto.org/docs/get-started/ "Quarto Download").
  - Download the installer for Windows.
  - Open a terminal to verify installation:
    - To launch the `Command Prompt`, type `cmd` in the Windows search field and press `enter`.
    - To launch `PowerShell`, type `PowerShell` in the search field and press `enter`.
    - In RStudio, click on the *Terminal* tab in the Console pane.
    - In VS Code, select `View > Terminal` or use the keyboard shortcut ``Ctrl + ` ` `
  - Confirm installation with:

```{.Powershell}
quarto --version
```
  
2. **Install RStudio and VS Code** (if they are not already installed on your machine).
3. **Install TeX Live**
  - Install the latest yearly release of TeX Live *or proceed to Step 4 if Tex Live is already installed*.
  - Please note that TeX Live is updated yearly, which requires full installation as each new release drops. The reason for this annual release schedule is to create a reliable and stable baseline for users with package updates synchronized with the core software updates.
  - Download [TeX Live](https://tug.org/texlive/ "TeX Live Download").
  - During install, ensure `tlmgr` (TeX Live Manager) is included.
  - Verify installation by running the following in a terminal:

```{.Powershell}
tlmgr --version
```

4. **Update TeX Live**
  - If Tex Live has already been installed on your machine, update it. In a terminal, run the following to  update `tlmgr` and to update all LaTeX packages:
  
```{.Powershell}
tlmgr update --self --all
```

5. **Verify PATH Setup**
  - Confirm `quarto --version`, `R --version`, and `python --version` all work in your terminal.
  - If any command fails, it means the software is not registered on your PATH. Re‑run the installer or update your PATH environment variable so the executables are discoverable. Guidance on how to do this can be available.

6. **Install Quarto Extensions**
  - Quarto Extensions are managed locally at the project level, so it will be necessary to add relevant extensions to each Quarto project folder rather than relying on a global installation. The reason for this is the need to ensure **reproducibility** and **portability** when projects are under version control. Bundling extensions with the project allows other users to confidently fork and clone projects without fear that they will not compile because the user lacks the specific version of the extension source code used as the project was developed (and maintained).
  - Therefore, projects may already have an `_extensions` folder with the required files. In that case, it may not be necessary to install extensions.
  - Adding new extensions to a project requires opening a terminal.
  - Identify and copy the file path to your projects root folder, which is the main directory that contains all of your project's files and subfolders, including the `.qmd` file.
  - In a terminal, type `cd` followed by the file path. For readability, the following is an example file path. You should replace it with the actual path on your system.

```{.Powershell}
cd C:\Users\bwing\OneDrive - ...\Documents\Repos - Quarto\Quarto Guide
```
  
  - From the project's root directory, run:
  
```{.Powershell}
quarto install repo-owner/repo-name
```
  
  - This will automatically create an `_extensions` folder if one does not already exist. Extensions are installed into a `quarto-ext/` folder at the project root. Quarto creates this folder automatically if it doesn't exist. In certain cases, a repository's `README.md` will instruct you to use `add` instead of `install`. Both will work, but `install` is the current convention because it clearly indicates that the operation is installing an extension.

---

## Chunk Options Cross-walk

### Chunk Options in R Markdown and Quarto Documents

Quarto moves chunk options out of the header and into the body of the chunk using YAML‑style comments (#|). This design makes options more consistent across languages and easier to read, especially when chunks get complex. The chunk label itself is unchanged, but execution and visibility options (eval, echo, include) are now written line‑by‑line inside the chunk.

| Purpose                 | R Markdown Syntax (chunk header) | Quarto Syntax (inside chunk)|
|-------------------------|----------------------------------|-----------------------------|
| Suppress code echo      | ````{r, echo=FALSE}````          | #\| echo: false             |
| Hide warnings           | ````{r, warning=FALSE}````       | #\| warning: false          |
| Hide messages           | ````{r, message=FALSE}````       | #\| message: false          |
| Hide results            | ````{r, results="hide"}````      | #\| results: hide           |
| Do not run code         | ````{r, eval=FALSE}````          | #\| eval: false             |
| Source Child Doc        | ````{r, child="insert path"}```` | #\| child: "insert path"    |
| Run but don’t include   | ````{r, include=FALSE}````       | #\| include: false          |
| Set fig width/height    | ````{r, fig.width=6,````         | #\| fig-width: 6            |
|                         |     ````fig.height=4}````        | #\| fig-height: 4           |
| Cache results           | ````{r, cache=TRUE}````          | #\| cache: true             |
| Chunk label             | ````{r my-chunk}````             | #\| label: my-chunk         |
| Parameters              | `params$start_date`              | `params.start_date`         |

### Global vs Local Chunk Options

In both R Markdown and Quarto, chunk options can be set globally (applied to all chunks) or locally (applied to a single chunk). Global options are useful for enforcing consistent defaults across a report, while local options let you override those defaults when needed. Quarto preserves this distinction, but the syntax shifts slightly.

| Scope   | R Markdown Syntax | Quarto Syntax |
|---------|-------------------|---------------|
| **Global** (all chunks) | In a setup chunk:<br>`knitr::opts_chunk$set(echo = FALSE, warning = FALSE)` | Same call in a setup chunk:<br>`knitr::opts_chunk$set(echo = FALSE, warning = FALSE)` |
| **Local** (single chunk) | ```{r my-chunk, echo=FALSE, warning=FALSE}` | ```{r my-chunk}`<br>`#| echo: false`<br>`#| warning: false` |


#### Global R Markdown

::: {.callout-tip #global-rmd-r-setup-callout}
````md
\```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
\```
````
:::

#### Global Quarto

::: {.callout-tip #global-qmd-r-setup-callout}
````md
\```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
\```
````
:::

#### Local R Markdown

::: {.callout-tip #local-rmd-r-setup-callout}
````md
\```{r summary, echo=TRUE}
summary(mtcars)
\```
````
:::

#### Local Quarto

::: {.callout-tip #local-qmd-r-setup-callout}
````md
\```{r summary}
#| echo: true
summary(mtcars)
\```
````
:::

## Callout Blocks in Quarto

Quarto supports several types of callout blocks to highlight information. These are built‑in and require no extensions.

### Basic Syntax

Callouts use the `:::` fence with a class. For example:

::: {.callout-note}
### Note
This is a note callout.
:::

### Types of Callouts

Below are examples of the main callout types. Each shows the source code and the rendered output.

#### Note - General Information

````md
::: {.callout-note}
### Note
This is a note callout.
:::
````

:::{.callout-note}
### Note
This is a note callout.
:::

#### Tip - Helpful Hints

````md
::: {.callout-tip}
### Tip
This is a tip callout.
:::
````

::: {.callout-tip}
### Tip
This is a tip callout.
:::

#### Warning - Advice About Serious Issues

````md
::: {.callout-warning}
### Warning
This is a warning callout.
:::
````

::: {.callout-warning}
### Warning
This is a warning callout.
:::

#### Caution - Cautionary Advice About Potential Problems

````md
::: {.callout-caution}
### Caution
This is a caution callout.
:::
````

::: {.callout-caution}
### Caution
This is a caution callout.
:::

#### Important - Critical Information

````md
::: {.callout-important}
### Important
This is an important callout.
:::
````

::: {.callout-important}
### Important
This is an important callout.
:::

Overall, note that callouts should be used for concise highlights. Do not place long lines of code in callout boxes because they do not break across pages by default.

````md
::: {#tip-cross-reference-example .callout-tip}
## Cross-Referencing a Tip

Add an ID starting with `#tip-` to reference a tip.
:::

See @tip-cross-reference-example...
````

This renders as:

::: {#tip-cross-reference-example .callout-tip}
## Cross-Referencing a Tip

Add an ID starting with `#tip-` to reference a tip.
:::

See @tip-cross-reference-example...

Each callout type has a specific prefix for this purpose.

| Callout Type | Prefix |
|--------------|--------|
| note         | #nte-  |
| tip          | #tip-  |
| warning      | #wrn-  |
| important    | #imp-  |
| caution      | #cau-  |


### Cross-Referencing Callouts

To cross-reference a callout, add an ID attribute starting with the appropriate callout prefix. This allows one to reference a callout using the usual @ syntax.

### Customization

- Titles can be added with an `title="..."` attribute to override the heading.
- Callouts can contain text, code blocks, tables, or inline icons.
- Icons and colors vary by type. 

### Callout Appearance Styles

Quarto supports three built‑in styles for callouts:

- **default** – full styling with border, background, and icon  
- **simple** – lighter styling with reduced border and background  
- **minimal** – text‑only, no border or background

#### Local (Per Callout)

You can set the style as an attribute on a single block:

````md
::: {.callout-tip appearance="default"}
### Default Tip
This callout uses the default style.
:::
````
::: {.callout-tip appearance="default"}
### Default Tip
This callout uses the default style.
:::

````md
::: {.callout-tip appearance="simple"}
### Simple Tip
This callout uses the simple style.
:::
````

::: {.callout-tip appearance="simple"}
### Simple Tip
This callout uses the simple style.
:::

````md
::: {.callout-tip appearance="minimal"}
### Minimal Tip
This callout uses the minimal style.
:::
````

::: {.callout-tip appearance="minimal"}
### Minimal Tip
This callout uses the minimal style.
:::

#### Global (All Callouts)

To apply a style to all callouts in a document, set the option in the YAML front matter:

````md
```yaml
format:
  html:
    callout-appearance: simple
  pdf:
    callout-appearance: minimal
```
````
### Icons in Callouts

By default, callouts include an icon. You can disable icons globally or locally.

#### Local (Per Callout)

Add icon=false to the block attributes:

````md
::: {.callout-note icon=false}
### Note Without Icon
This callout has no icon.
:::
````

#### Global (All Callouts)

Set the option in the YAML front matter:

````md
---
format:
  html:
    callout-icon: false
  pdf:
    callout-icon: false
---
````

#### Enhancing Callouts With Additional Icons

Callouts can be enhanced with icons for emphasis. For example, you may insert Font Awesome icons inline to visually distinguish tips, warnings, or highlights. See the [Font Awesome section](#font-awesome) later in this guide for detailed instructions on customizing icons (color, size, alignment) in both PDF and HTML outputs.

## Package Compatibility Notes

### **Bookdown Features**

  - Quarto subsumes bookdown features by natively supporting multi-file books, cross-references, citations, and output formats without requiring the bookdown package.
  - Quarto projects use a unified YAML project file (`_quarto.yml`) to define book structure, replacing bookdown's `_bookdown.yml`.
  - Quarto supports child documents and parameterized reports similarly to bookdown but with enhanced syntax and cross-language support.
  - Quarto's native support for extensions and filters replaces many bookdown-specific features, providing more flexibility and maintainability.

### **Additional common package dependencies**
  - TBD

## Project Structures

### Parent/Child Document Model

- Quarto supports `include:` and `import:` directives, which means that our child ETL documents may still be source.

### Parameters for Quarterly vs Annual Reports

### Conditional Child Docs

#### Project Configuration File

::: {.callout-tip #yml-file-pattern-callout title="`_quarto.yml`"}
````md
```yaml
project:
  type: book

book:
  title: "YERE Program Report"
  chapters:
    - index.qmd
    - 01-YERE-Program-Periodic-Report.qmd
```
````
:::

#### Front Matter and Setup

::: {.callout-tip #index-rmd-pattern-callout title="`index.Qmd`"}
````md
```yaml
---
title: "YERE Program Report"
subtitle: "Community-Based Access to Care"
params:
  start_date: !expr lubridate::ymd("2025-04-01")
  end_date: !expr lubridate::ymd("2025-06-30")
  hospital_system: "SSM"
---
```
````
:::

#### R Setup Chunk

::: {.callout-tip #yere-r-setup-example-pattern-callout title="R Set Up Chunk"}
````md
```{r r-setup}
#| include: FALSE
#| eval: FALSE

time_difference <- params$end_date - params$start_date

if (time_difference < 364) {
  report_type <- "Quarterly"
} else {
  report_type <- "Annual"
}

# Define evaluation flags
eval_qtr <- (report_type == "Quarterly")
eval_annual <- (report_type == "Annual")
eval_general <- TRUE

# Fiscal period string
fiscal_period <- ifelse(report_type == "Quarterly", "quarters", "years")
```
````
:::

#### Parent File

::: {.callout-tip #yere-parent-file-example-pattern-callout title="01-YERE-Program-Periodic-Report.qmd"}
```md
# YERE Program Report
```
````{r, eval=FALSE}
```{r}
#| label: yere-section-geography
#| child: "Sections/YERE-Section-Geography.qmd"
#| eval: eval_general
```

```{r}
#| label: yere-section-quarterly
#| child: "Sections/YERE-Section-Quarterly.qmd"
#| eval: eval_annual
```
````
:::


### Multi-File Book Projects (e.g., YERE)

### Example YAML Setups

## Tables

We have used the kableExtra package to format tables in our report projects because kableExtra excels at implementing LaTeX features. The good news is that kableExtra works in Quarto just as it did in R Markdown projects, since Quarto still uses knitr for R execution. However, kableExtra tend to be fragile when rendering to Word .docx files, and the styling tends to be lost. This is where the flextable package offers some benefit because flextable was written to support Word output. The good news is that we don't have to choose between the two. Instead, we can adopt a hybrid strategy for table rendering: conditionally use kableExtra for PDF and flextable for Word output.

### kableExtra and flextable Function Cross-Walk

| Task / Feature        | kableExtra (PDF/HTML)                          | flextable (Word/PPT)                       |
|-----------------------|------------------------------------------------|--------------------------------------------|
| Basic table           | `kable(df)`                                    | `flextable(df)`                            |
| Add header row        | `add_header_above()`                           | `add_header_row()`                         |
| Bold header row       | `row_spec(0, bold=TRUE)`                       | `bold(part="header")`                      |
| Last row styling      | `row_spec(nrow, bold=TRUE, bg="#2c7fb8")`      | `bg(i=nrow, bg="#2c7fb8"); color(i=nrow, color="white")`                      |
| Group cells           | `collapse_rows(columns=1)`                     | `merge_v(j=1)`                             |
| Style header          | `kable_styling(latex_options="striped")`       | `theme_vanilla()` or `bold(part="header")` |
| Cell background color | `cell_spec(..., background="lightblue")`       | `bg(i=1, j=1, bg="lightblue")`             |
| Column width          | `column_spec(2, width="5cm")`                  | `width(j=2, width=2)`                      |
| Footnotes             | `footnote(general="Note...")`                  | `add_footer_lines("Note...")`              |
| Alignment             | `align="c"` in `kable()`                       | `align(j=2, align="center")`               |
| Caption               | `caption=` in `kable()`                        | `flextable::set_caption()`               |

### Conditional Routing Example

Because kableExtra and flextable are used to format and style tables, the data preparation steps may remain the same. Once data has been transformed for use in a table and assigned to a tibble, we would then pass the object to a function that 

````{r, eval=FALSE}
```r
df <- data.frame(
  Name = c("Alice", "Bob"),
  Score = c(95, 88)
)

if (knitr::is_latex_output()) {
  # PDF path
  df %>%
    knitr::kable("latex", booktabs = TRUE) %>%
    kableExtra::kable_styling(latex_options = "striped")
} else if (knitr::is_html_output()) {
  # HTML path
  df %>%
    knitr::kable("html") %>%
    kableExtra::kable_styling(full_width = FALSE)
} else if (knitr::is_word_output()) {
  # Word path
  flextable::flextable(df) %>%
    flextable::theme_vanilla()
}
```
````

- Conditional routing example chunk
- Helper function template (like make_nested_header_table())
- Notes on stakeholder customization (n + % columns, narrative fallbacks)
- Icons in PDF vs Word outputs

### Table Toolkit

**Nested Table Header Example**

An illustration of code that might be adopted to conditionally pipe tables requiring nested headers follows:

````{r, eval=FALSE}
```{r make-nested-header-table-template}
#| echo: false
#| warning: false
#| eval: false

make_nested_header_table <- function(df, col_labels, caption_text) {
  if (knitr::is_latex_output()) {
    df %>%
      knitr::kable(
        caption = caption_text,
        booktabs = TRUE,
        escape = FALSE,
        format.args = list(big.mark = ","),
        linesep = "",
        col.names = linebreak(col_labels, align = "c"),
        align = c("l", rep("c", ncol(df) - 1))
      ) %>%
      collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
      kableExtra::kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
      row_spec(0, bold = TRUE) %>%
      row_spec(nrow(df), bold = TRUE, color = "white", background = "#2c7fb8") %>%
      add_header_above(
        c(" " = 2, "Two Week" = 2, "Thirty Day" = 2, "Three Month" = 2, "Six Month" = 2),
        bold = TRUE
      )
  } else if (knitr::is_word_output()) {
    flextable::flextable(df) %>%
      flextable::set_header_labels(.labels = col_labels) %>%
      flextable::add_header_row(
        values = c("", "Two Week", "Thirty Day", "Three Month", "Six Month"),
        colwidths = c(2, 2, 2, 2, 2)
      ) %>%
      flextable::bold(part = "header") %>%
      flextable::bg(i = nrow(df), bg = "#2c7fb8") %>%
      flextable::color(i = nrow(df), color = "white") %>%
      flextable::autofit() %>%
      flextable::set_caption(caption_text)
  }
}
```
````


The code to use this in a program report chunk:

::: {.callout-tip #nested-header-table-application-pattern-callout}
````{r, eval=FALSE}
```{r make-nested-header-table-application}
#| echo: false
#| warning: false
#| eval: false

make_nested_header_table(
  epicc_tbl_agency_eligible_part_three_fps,
  colnames_agency_eligible_part_three_fps_string,
  paste0(
    "Client Program Participation Over Time by Agency, ",
    current_reporting_fiscal_period
    )
  )
```
````
:::

## Output Formats

- Rendering to PDF and Word simultaneously
- Critic markup for PDF review
- Limitations of Word output (icons, LaTeX features)

## Migration Guidance

- Converting .Rmd → .qmd
- Updating chunk options
- Handling parameters and child documents
- Strategy for phased adoption

In the next section, we’ll look at look at examples of specific extensions that are potentially relevant for our reporting needs.

---

## Quarto Extensions

## Critic Markup

The Critic Markup extension offers specialized, Markdown-based syntax to denote edits and revisions when submitting report drafts for review by internal agency stakeholders. The syntax is quick and easy to add and remove from the document. This section shows the Critic Markup syntax verbatim, alongside rendered examples.

::: {.callout-note}
### Note
Although the Critic Markup extension installs into `quarto-ext/critic-markup/`, you must list `critic-markup` (*without `.lua`) under `filters:` and not under `extensions:` as one might expect. Quarto reads the extension's `_extension.yml` file and registers the `Lua` filter automatically, which rewrites the syntax into LaTeX macros (e.g., highlighting, colored text for additions and deletions). In HTML output, the Critic Markup extension uses CSS/JS for styling, so listing the extension as `mloubout/critic-markup` under `extensions:` would result in no styling being applied to the PDF.
:::

## Highlighting and Commenting

It is quite simple to highlight a block of text by wrapping it as shown:

`{==Highlighted text==}` {==Highlighted text==}

Comments may also be added, either in conjuction with highlighting or without:

`{== Highlighted text with a comment==}{>>First comment<<}` {== Highlighted text with a comment==}{>>First comment<<}

`{>>Single comment <<}` {>>Single comment <<}

## Showing the Diff: Addition, Deletion, and Replacement of Text

Critic Markup allows for diff-style markup by enabling us to wrap text to signify additions and deletions from the document:

`{++This is added text++}` {++This is added text++}

`{++This is added text with some removed++}{-- removed--}` {++This is added text with some removed++}{-- removed--}

`{--This is removed text--}` {--This is removed text--}

`{--This is removed text with some added--}{++added ++}` {--This is removed text with some added--}{++added ++}

`{~~This is original text~> this is the replacement~~}` {~~This is original text~> this is the replacement~~}

---

## Font Awesome

Quarto may also be extended to include the ability to add scalable, vector-based icons to enhance the text. The vast majority of icon sets are designed for use exclusively by web developers using HTML, which means that the extensions require quite a bit of work to configure them for use with LaTeX. Fortunately, a Font Awesome extension is available to allow us to use the Font Awesome (FA) library of icons in PDF reports.

::: {.callout-note}
### Note
Currently, the Font Awesome extension is configured to work with Font Awesome 5 free icons only when compiling to PDF and to otherwise use Font Awesome 6 icons when compiling to HTML. However, because Pandoc scans for both Font Awesome 5 and 6, the result is that compiling to PDF fails unless the extension is modified to force using FA 5 exclusively. I have modified the `fontawesome.lua` file to do this. Issues in the repository indicate that the developers plan to add LaTeX support for Font Awesome 6 and 7 libraries in the future. We can fork the repo to ensure that we are aware of updates as they are merged with the main branch.
:::

As this section will demonstrate, FA icons are easy to customize.

### Basic Usage

Without changing the font-size or color, FA icons will be solid and will compile in black and white:

- \faBullhorn
- \faCheckCircle
- \faClipboard

### Adding Color

All FA icons may be filled with color:

- \textcolor{Red}{\faChartLine} 14 program outcomes improved
- \textcolor{Blue}{\faUsers} Community participation increased by 45%
- \textcolor{ForestGreen}{\faLeaf} 3 sustainability initiatives expanded

### Changing Size and Color of Inline Icons

In addition to adding color, FA icons may also be customized to increase the size of the icons relative to the text:

```{=latex}
\begin{tabular}{ll}
  \textbf{Color only} & \textcolor{Blue}{\faUsers} \\
  \textbf{Size 2} & \textcolor{Blue}{\scalebox{2}{\faUsers}} \\
  \textbf{Size 3} & \textcolor{Blue}{\scalebox{3}{\faUsers}} \\
  \textbf{Size 4} & \textcolor{Blue}{\scalebox{4}{\faUsers}}
\end{tabular}
```

### Callout Note Box With Inline Icon

Markdown callout blocks are built into Quarto. These may be used to highlight information (notes, tips, warnings, examples). Notably, icons may be added to these blocks:

::: {.callout-note}
### Highlight
\textcolor{Orange}{\faLightbulb} 25 grant applications were submitted
:::

### Callout-Style Note Box With Icon Aligned Left and Not Inline

It is possible to manually construct note boxes that resemble callout blocks, with the advantage that these allow for fine-grained control over layout (icon alignment, spacing, colors):

```{=latex}
\begin{tcolorbox}[colback=gray!10,colframe=gray!50]
  \begin{tabular}{@{}l p{0.8\textwidth}@{}}
    \parbox[t]{0.1\textwidth}{\centering
      \raisebox{-\height+\depth+0.25ex}{\scalebox{3}{\textcolor{Bittersweet}{\faBullhorn}}}
    } &
    \parbox[t]{\linewidth}{
      2 new workflows were adopted\\
      32 new staff were trained\\
      450 assessments were completed
    }
  \end{tabular}
\end{tcolorbox}
```

### Icon and Text Side by Side (No Wrapping)

Icons and text may be typeset side by side without wrapping text around the icon even when a block of text is lengthy:

```{=latex}
\begin{minipage}[t]{0.15\textwidth}
  \raisebox{\dimexpr-\height+\depth}{%
    \scalebox{5}{\textcolor{TealBlue}{\faChartBar}}%
  }
\end{minipage}%
\begin{minipage}[t]{0.8\textwidth}
  \lipsum[1]
\end{minipage}
```

### Icon and Text Side by Side (With Wrapping)

It is also possible to wrap text around icons to mimic the newspaper style:

```{=latex}
\begin{wrapfigure}{l}{0.15\textwidth} % l = left side, width of box
  \centering
  \scalebox{5}{\textcolor{RoyalBlue}{\faAmbulance}}
\end{wrapfigure}
\lipsum[1]
```

### Facet Row (Icons Plus Caption)

By combining a tabular environment with `\parbox` commands, which formats text in an invisible container in paragraph mode to enable text wrapping and justification, it is possible to create a row of "facets" with a big icon on top with a text caption underneath each one:

```{=latex}
\begin{center}
\begin{tabular}{ccc}
  \parbox[t]{0.25\textwidth}{\centering
    \scalebox{4}{\textcolor{Cerulean}{\faChartArea}}\\[0.5ex]
    5 metrics were approved and adopted
  } &
  \parbox[t]{0.25\textwidth}{\centering
    \scalebox{4}{\textcolor{CornflowerBlue}{\faChurch}}\\[0.5ex]
    5 new churches were onboarded
  } &
  \parbox[t]{0.25\textwidth}{\centering
    \scalebox{4}{\textcolor{Mulberry}{\faChartPie}}\\[0.5ex]
    0 exploding pie charts were used in reports (and never will be)
  }
\end{tabular}
\end{center}
```

### Icons as Bullet Marker

Last, it is possible to create lists with icons in place of bullets:

```{=latex}
\begin{itemize}
  \item[\textcolor{JungleGreen}{\faDatabase}] FAMCare database went live
  \item[\textcolor{Blue}{\faClinicMedical}] 1 new hospital partner was engaged
  \item[\textcolor{Orchid}{\faBed}] 2 shelter beds were allocated
\end{itemize}
```

---

## Bootstrap Icons

Another free, open-source library of icons is Bootstrap Icons (bsicons). Bootstrap Icons are  designed to be scalable and easy to style ... in HTML. Since these are saved as SVG icons, styling in LaTeX is much more challenging than with Font Awesome icons. It just isn't worth our time and effort to customize bsicons with different colors, so these icons will only render in black. However, they are easy to insert into Quarto documents and will compile to PDF now that I have customized the `Lua` script, so it may be worth perusing the library to see if there are icons that would be a good complement to the Font Awesome 5 library.

### Basic Syntax

I have modified the .Lua filter in this extension to make it easy for us to render bsicons to PDF with the following basic syntax:

- Alarm {{< bi alarm-fill size=2 >}}
- Archive Filled {{< bi archive-fill size=2 >}}
- Award {{< bi award size=2 >}}
- Bar Chart Steps {{< bi bar-chart-steps size=2 >}}
- Bell Filled {{< bi bell-fill size=2 >}}
- Book, Half Filled {{< bi book-half size=2 >}}
- Bullseye {{< bi bullseye size=2 >}}
- Card With Checklist {{< bi card-checklist size=2 >}}
- Chat With Dots {{< bi chat-dots size=2 >}}
- Heart With Pulse Filled {{< bi heart-pulse-fill size=2 >}}
- House {{< bi house size=2 >}}
- Key {{< bi key size=2 >}}
- Lightning {{< bi lightning size=2 >}}
- Octagon Filled {{< bi octagon-fill size=2 >}}
- Person Badge Filled {{< bi person-badge-fill size=2 >}}
- Rocket Taking Off {{< bi rocket-takeoff size=2 >}}
- Sliders {{< bi sliders size=2 >}}
- Trophy Filled {{< bi trophy-fill size=2 >}}
- Trashcan {{< bi trash size=2 >}}
- X-Diamond Filled {{< bi x-diamond-fill size=2 >}}

### Resizing the Icons

These icons may be resized, as illustrated by the following renderings of a Bar Chart With X-Axis Line:

- Size 1 {{< bi bar-chart-line size=1 >}}
- Size 2 {{< bi bar-chart-line size=2 >}}
- Size 3 {{< bi bar-chart-line size=3 >}}
- Size 4 {{< bi bar-chart-line size=4 >}}

### Multiple Icon Versions

A number of icons come in multiple versions as this non-exhaustive list of house icons illustrates:

- House {{< bi house size=2 >}}
- House With Addition Sign {{< bi house-add size=2 >}}
- House Filled With Addition Sign {{< bi house-add-fill size=2 >}}
- House With Check {{< bi house-check size=2 >}}
- House Filled With Check {{< bi house-check-fill size=2 >}}
- House With Door {{< bi house-door size=2 >}}
- House Filled With Door {{< bi house-door-fill size=2 >}}
- House With Heart {{< bi house-heart size=2 >}}
- Houses {{< bi houses size=2 >}}

---

## Appendix

- Reference snippets (global chunk options, table toolkit)
- Links to Quarto documentation